Optimal Pricing and Informal Sharing: \\ Evidence from Piped Water in Manila

### 2018-09-25 phil_paper/jmp_wjv_v2.tex
add out of sample prediction

\begin{figure}
\caption{Comparing Predicted and Observed Water Use}\label{figure:predictedwateruse}
\begin{center}
\includegraphics[scale=.77]{tables/hist_both.pdf}
\end{center}
\end{figure}

Figure~\ref{figure:predictedwateruse} compares the predicted histogram of consumption with observed values.  Vertical red lines indicate the price thresholds in the tariff schedule.  By roughly matching the smoothness of the distributions especially across price thresholds, this model improves on standard models in the literature, which often predict excessive bunching at thresholds (\cite{mcrae2014infrastructure}, \cite{olmstead2009reduced}, \cite{szabo2015value}).  The model has difficultly rationalizing low levels of consumption likely because households face zero prices below this threshold.  Since preference and consumption shocks are assumed to be normally distributed, the model predicts some negative consumption values.  In the counterfactual exercises, I treat these negative realizations as zero consumption choices.  Likewise, the model also cannot account for bunching at zero (likely capturing households that go on vacation).  The model also predicts greater bunching around the upper price threshold despite little observed bunching in the data as large users tend to ignore this smaller jump in price.

@@

With the full set of household preferences, the simulation routine first groups households into random sets of three neighbors within each of the 599 wards.  The routine then orders households in the sequential game according to their utilities from using an individual connection.  This assumption captures the intuition that households who would have connected in the absence of any sharing are likely to be the first to take-up a new connection.  Random draws of the preference and consumption shocks, $\eta$ and $\epsilon$, are assigned to each household so that households can compute expected utilities over \input{tables/sim_time_periods}months for all possible water source choices.  These shocks are sampled from a multivariate normal distribution taking into account the estimated correlation between sharing partners presented in Table~\ref{table:correstimates}.  
%\footnote{ Future versions will not condition on observed choices in this way, and instead sample households across sharing relationships. }

>>>>

%is its inability to match zero consumption values observed in the data.

\begin{table}
\caption{Model Performance}\label{table:modelperformance}
\input{tables/sample_prediction}
\end{table}

Table~\ref{table:modelperformance} compares predicted consumption from the model to observed consumption in the data.  This exercise finds negligible differences across average consumption and the variance of consumption for connections according to sharing status.  Closely matching observed patterns, this model improves on previous approaches to demand estimation likely by recovering a separate fixed preference term for each connection as well as rich heterogeneity in the variance terms (\cite{szabo2015value}).

%\begin{figure}
%\caption{Comparing Predicted and Observed Water Use}\label{figure:predictedwateruse}
%\begin{center}
%\includegraphics[scale=.77]{tables/hist_both.pdf}
%\end{center}
%\end{figure}

%Figure~\ref{figure:predictedwateruse} compares the predicted histogram of consumption with observed values.  Vertical red lines indicate the price thresholds in the tariff schedule.  By roughly matching the smoothness of the distributions especially across price thresholds, this model improves on standard models in the literature, which often predict excessive bunching at thresholds (\cite{mcrae2014infrastructure}, \cite{olmstead2009reduced}, \cite{szabo2015value}).  The model has difficultly rationalizing low levels of consumption likely because households face zero prices below this threshold.  Since preference and consumption shocks are assumed to be normally distributed, the model predicts some negative consumption values.  In the counterfactual exercises, I treat these negative realizations as zero consumption choices.  Likewise, the model also cannot account for bunching at zero (likely capturing households that go on vacation).  The model also predicts greater bunching around the upper price threshold despite little observed bunching in the data as large users tend to ignore this smaller jump in price.

@@

With the full set of household preferences, the simulation routine first groups households into random sets of three neighbors within each of the 599 wards.  The routine then orders households in the sequential game according to their utilities from using an individual connection.  This assumption captures the intuition that households who would have connected in the absence of any sharing are likely to be the first to take-up a new connection.  Random draws of the preference and consumption shocks, $\eta$ and $\epsilon$, are assigned to each household so that households can compute expected utilities over \input{tables/sim_time_periods}months for all possible water source choices.  These shocks are sampled from a multivariate normal distribution taking into account the estimated correlation between sharing partners presented in Table~\ref{table:correstimates}.  One limitation of the model is that the linear demand structure with normal error terms also predicts occasional negative consumption values.  Those values are rounded to zero for calculating household utility.

%\footnote{ Future versions will not condition on observed choices in this way, and instead sample households across sharing relationships. }
